---
title: "Data Preparation Process"
author: "Wilmer Uruchi"
date: "12/09/2018"
output:
  html_document:
    fig_height: 4
    highlight: pygments
    theme: spacelab
  pdf_document: 
    keep_tex: yes
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Building the DataSet

First, visualize the current dataset:

```{r visualize Boston}
load("Boston2017.RData")

```

Using dplyr for some data transformation:
```{r load_libraries1, message=FALSE}
library(dplyr)
library(lubridate)
library(stringr)
library(ggplot2)
```

```{r load_libraries2, message=FALSE}
library("FactoMineR")
```

I am going to build the dataset using the imported dataset marathon_results_2017, and selecting the participants between 20 and 60 years old, inclusive, to avoid outliers.

```{r new dataset}
Boston2017SubSet <- marathon_results_2017 %>%
  filter(Age >= 20, Age <= 60)
```

Analyze Data Types:
```{r}
str(Boston2017SubSet)

```

Now, I am going to transform some fields so I can use them in the PCA.
```{r transform}
Boston2017SubSet <- Boston2017SubSet %>%
  mutate(X10ktotal = as.numeric(X10K), PaceTotal = as.numeric(Pace), 
         X5kseconds = as.numeric(difftime(as.POSIXct(X5K, format = '%H:%M:%S'), as.POSIXct('00:00:00', format = '%H:%M:%S'), units = 'secs')),
         X10kseconds = as.numeric(difftime(as.POSIXct(X10K, format = '%H:%M:%S'), as.POSIXct('00:00:00', format = '%H:%M:%S'), units = 'secs')),
         PaceSeconds = as.numeric(difftime(as.POSIXct(Pace, format = '%H:%M:%S'), as.POSIXct('00:00:00', format = '%H:%M:%S'), units = 'secs')),
         OfficialTimeSeconds = as.numeric(difftime(as.POSIXct(Official.Time, format = '%H:%M:%S'), as.POSIXct('00:00:00', format = '%H:%M:%S'), units = 'secs')),
         BibNumber = str_extract(Bib, "[F]"))
```

To achieve a homogeneous group, we are not going to include all the professionals into the model since they are expected to have results that are about 3 standard deviations from the mean. 

```{r}
Boston2017NoPro <- Boston2017SubSet %>%
  filter(is.na(BibNumber) == TRUE, is.na(OfficialTimeSeconds) == FALSE, is.na(X5kseconds) == FALSE,
         is.na(X10kseconds) == FALSE)
```

Another transformation:

```{r}
Boston2017NoPro <- Boston2017NoPro %>%
  mutate(BibN = as.numeric(as.character(Bib)))
Boston2017NoPro <- Boston2017NoPro %>%
  mutate(GenderN = ifelse(M.F == "M",1,2))
```


## Doing PCA

Now lets look at the PCA of the data collected:

```{r pressure, echo=FALSE}
Boston2017NoPro.PCA<-Boston2017NoPro[, c("X5kseconds", "Age", "BibN", "GenderN")]
res<-PCA(Boston2017NoPro.PCA , scale.unit=TRUE, ncp=5, graph = FALSE)
plot.PCA(res, axes=c(1, 2), choix="ind", habillage="none", col.ind="black", 
  col.ind.sup="blue", col.quali="magenta", label=c("ind", "ind.sup", "quali"),
  new.plot=TRUE)
plot.PCA(res, axes=c(1, 2), choix="var", new.plot=TRUE, col.var="black", 
  col.quanti.sup="blue", label=c("var", "quanti.sup"), lim.cos2.var=0)
summary(res, nb.dec = 3, nbelements=10, nbind = 10, ncp = 3, file="")
```

It looks like these are the variable that should be included in the construction of the model:
```{r model-construction}
RegModel.2 <- lm(OfficialTimeSeconds~Age+X5kseconds+BibN+GenderN, 
  data=Boston2017NoPro)
summary(RegModel.2)
```

Distribution of 5k times:
```{r}
ggplot(data = Boston2017NoPro, aes(x = X5kseconds)) +
  geom_histogram()
Boston2017NoPro %>%
  summarise(mean_ar = mean(X5kseconds), median = median(X5kseconds), sd_ar = sd(X5kseconds), n = n(), max = max(X5kseconds), min = min(X5kseconds), sd_ar = sd(X5kseconds), iqr_ar = IQR(X5kseconds))
```

```{r}
ggplot(data = Boston2017NoPro, aes(x = OfficialTimeSeconds)) +
  geom_histogram(bins = 30) +
  labs(y = "runners", x = "time in seconds")

ggplot(data = Boston2017NoPro, aes(x = OfficialTimeSeconds, y = BibN)) +
  geom_line()

Boston2017NoPro %>%
  summarise(mean_ar = mean(OfficialTimeSeconds), median = median(OfficialTimeSeconds), sd_ar = sd(OfficialTimeSeconds), n = n(), max = max(X5kseconds), min = min(OfficialTimeSeconds), sd_ar = sd(OfficialTimeSeconds), iqr_ar = IQR(OfficialTimeSeconds))
```


```{r}
ggplot(regularSim, aes(Num)) + 
  geom_line(aes(y = OfficialTime, colour = "Official")) + 
  geom_line(aes(y = Result, colour = "Result")) +
  ggtitle("Results") +
  labs(y = 'time')

ggplot(regularSim, aes(x = OfficialTime)) +
  geom_histogram(bins =30)

ggplot(regularSim, aes(x = Result)) +
  geom_histogram(bins =30)
```

```{r}
R1Sim_Bib.Gender.Age.factors %>%
  filter(Gender == 2) %>%
  summarise(mean = mean(Result))

`R3Sim_Bib.Gender+Age.factors` %>%
  filter(Gender == 1) %>%
  summarise(mean = mean(Result))


meanR1 = mean(R1Sim_Bib.Gender.Age.factors$Result)
meanR2 = mean(`R2Sim_Bib.Gender.Age+factors`$Result)
meanR3 = mean(`R3Sim_Bib.Gender+Age.factors`$Result)
meanR4 = mean(`R4Sim_Bib.Gender+Age+factors`$Result)
meanR5 = mean(`R5Sim_Bib+Gender.Age.factors`$Result)
meanR6 = mean(`R6Sim_Bib+Gender.Age+factors`$Result)
meanR7 = mean(`R7Sim_Bib+Gender+Age.factors`$Result)
meanR8 = mean(`R8Sim_Bib+Gender+Age+factors`$Result)
meanR1
meanR2
meanR3
meanR4
meanR5
meanR6
meanR7
meanR8

e1 = (meanR2 - meanR1 + meanR4 - meanR3 + meanR6 - meanR5 + meanR8 - meanR7)/4
e1
e2 = (meanR3 + meanR4 + meanR7 + meanR8 - meanR1 - meanR2 - meanR5 - meanR6)/4
e2
e3 = (meanR5 + meanR6 + meanR7 + meanR8 - meanR1 - meanR2 - meanR3 - meanR4)/4
e3
```


```{r}
regularSim %>%
  summarise(mean_sim = mean(Result), mean_his = mean(OfficialTime), sd_sim = sd(Result), sd_his= sd(OfficialTime))

```

<!-- ```{r Export Data} -->
<!-- Boston2017EmpDistro <- Boston2017NoPro %>% -->
<!--   select(BibN, GenderN, Age, X5kseconds, OfficialTimeSeconds) -->
<!-- write.table(Boston2017EmpDistro, "EmpiricalData.txt", sep=",") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- Boston2017EmpDistro %>% -->
<!--   summarise(maxAge = max(Boston2017EmpDistro$Age), minAge = min(Boston2017EmpDistro$Age), minBib = min(Boston2017EmpDistro$BibN), maxBib = max(Boston2017EmpDistro$BibN)) -->
<!-- ``` -->



<!-- ```{r} -->
<!-- ggplot(data = log, aes(x = V1)) + -->
<!--   geom_histogram() -->

<!-- log %>% -->
<!--   summarise(mean_ar = mean(V1), median = median(V1), sd_ar = sd(V1), n = n(), max = max(V1), min = min(V1), var_ar = var(as.numeric(as.character(V1))), alpha = (mean(V1 * V1))/var_ar, beta = (mean_ar/var_ar)) -->

<!-- ``` -->

<!-- ```{r} -->
<!-- ggplot(data = regularSim, aes(x = Result)) + -->
<!--   geom_histogram() -->

<!-- regularSim %>% -->
<!--   summarise(mean_ar = mean(Result), median = median(Result), sd_ar = sd(Result), n = n(), max = max(Result), min = min(Result)) -->

<!-- ggplot(data = Boston2017NoPro, aes(x = OfficialTimeSeconds)) + -->
<!--   geom_histogram() -->

<!-- Boston2017NoPro %>% -->
<!--   summarise(mean_ar = mean(OfficialTimeSeconds), median = median(OfficialTimeSeconds), sd_ar = sd(OfficialTimeSeconds), n = n(), max = max(OfficialTimeSeconds), min = min(OfficialTimeSeconds)) -->

<!-- ``` -->

<!-- Analyzing Negative Age (20): -->
<!-- ```{r} -->
<!-- ggplot(data = ageNegativeSim, aes(x = Result)) + -->
<!--   geom_histogram() -->

<!-- ageNegativeSim %>% -->
<!--   summarise(mean_ar = mean(Result), median = median(Result), sd_ar = sd(Result), n = n(), max = max(Result), min = min(Result)) -->
<!-- ``` -->

<!-- Analyzing Positive Age (60): -->
<!-- ```{r} -->
<!-- ggplot(data = agePositiveSim, aes(x = Result)) + -->
<!--   geom_histogram() -->

<!-- agePositiveSim %>% -->
<!--   summarise(mean_ar = mean(Result), median = median(Result), sd_ar = sd(Result), n = n(), max = max(Result), min = min(Result)) -->
<!-- ``` -->

<!-- Analyzing Positive Gender (Male): -->
<!-- ```{r} -->
<!-- ggplot(data = genderPositiveSim, aes(x = Result)) + -->
<!--   geom_histogram() -->

<!-- genderPositiveSim %>% -->
<!--   summarise(mean_ar = mean(Result), median = median(Result), sd_ar = sd(Result), n = n(), max = max(Result), min = min(Result)) -->
<!-- ``` -->

<!-- Analyzing Negative Gender (Female): -->
<!-- ```{r} -->
<!-- ggplot(data = genderNegativeSim, aes(x = Result)) + -->
<!--   geom_histogram() -->

<!-- genderNegativeSim %>% -->
<!--   summarise(mean_ar = mean(Result), median = median(Result), sd_ar = sd(Result), n = n(), max = max(Result), min = min(Result)) -->
<!-- ``` -->

<!-- Analyzing Positive Bib (31437): -->
<!-- ```{r} -->
<!-- ggplot(data = bibPositiveSim, aes(x = Result)) + -->
<!--   geom_histogram() -->

<!-- bibPositiveSim %>% -->
<!--   summarise(mean_ar = mean(Result), median = median(Result), sd_ar = sd(Result), n = n(), max = max(Result), min = min(Result)) -->
<!-- ``` -->

<!-- Analyzing Negative Bib (3): -->
<!-- ```{r} -->
<!-- ggplot(data = `R8Sim_Bib+Gender+Age+factors`, aes(x = Result)) + -->
<!--   geom_histogram() -->

<!-- `R8Sim_Bib+Gender+Age+factors` %>% -->
<!--   summarise(mean_ar = mean(Result), median = median(Result), sd_ar = sd(Result), n = n(), max = max(Result), min = min(Result)) -->
<!-- ``` -->

